SI: Settlements / Pottery groups

### 2022-06-23 analysis/groups/groups.Rmd
analysis for pottery groups

```{r plot-kde}
kde %>% ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "KDE")
```

@@

```{r alt}
settlements2 %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, altitude)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Altitude (m)")
```

@@

```{r plot-slope}
slope %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Slope (°)")
```

@@

```{r plot-tpi}
tpi %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "TPI")
```

@@

```{r plot-hydro}
hydro %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^.")) %>% 
  ggplot(aes(period, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  labs(y = "Hydrology KDE")
```

@@

@@

```{r pca-plot}
pca_per %>% plot_nested_pca(var = "period")
```

>>>>

```{r plot-kde}
kde %>% 
  add_period_label() %>% 
  mutate(reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  add_period_label() %>% 
  ggplot(aes(period_label, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "KDE (Density of settlements per square km)", x = "Period")

# ggsave(here::here("groups_kde.svg"), width = 10, height = 5)
```

@@

```{r alt}
settlements2 %>% 
  mutate(reg = str_extract(id, "^."),
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  add_period_label() %>% 
  ggplot(aes(period_label, altitude)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Altitude (m)", x = "Period")

# ggsave(here::here("groups_altitude.svg"), width = 10, height = 5)
```

@@

```{r plot-slope}
slope %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^."),
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  add_period_label() %>% 
  ggplot(aes(period_label, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Slope (°)", x = "Period")

# ggsave(here::here("groups_slope.svg"), width = 10, height = 5)
```

@@

```{r plot-tpi}
tpi %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^."),
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  add_period_label() %>% 
  ggplot(aes(period_label, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "TPI", x = "Period")
```

```{r}
# ggsave(here::here("groups_tpi.svg"), width = 10, height = 5)
```

@@

```{r plot-hydro}
hydro %>% 
  left_join(settlements2, by = "id") %>% 
  mutate(reg = str_extract(id, "^."),
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  add_period_label() %>% 
  ggplot(aes(period_label, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Hydrology KDE", x = "Period")
```
```{r}
# ggsave(here::here("groups_hydro.svg"), width = 10, height = 5)
```

@@

```{r}
rm_dist_fin %>% 
  filter(variable == "rm_dist_chipped") %>% 
  add_period_label() %>% 
  mutate(reg = str_extract(id, "^."),
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>% 
  ggplot(aes(period_label, value)) +
  geom_violin() +
  geom_boxplot(alpha = 0.4, width = 0.2) +
  geom_jitter(alpha = 0.05, width = 0.4) +
  facet_wrap(vars(reg)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, hjust = 0)) +
  labs(y = "Distance to lithic raw. mat. (km)", x = "Period")

# ggsave(here::here("groups_distSI.svg"), width = 10, height = 5)
```

# Continuity

```{r continuity-calc}
cont <- set_base$period2 %>%
  select(-period_label, -chrono, - facet) %>%
  distinct(id, period) %>%
  mutate(pres = 1L) %>%
  pivot_wider(names_from = period, values_from = pres, values_fill = 0L) %>%
  pivot_longer(starts_with("p"), names_to = "period") %>%
  mutate(period = factor(period, levels = levels(set_base$period2$period))) %>%
  arrange(id, period) %>%
  group_by(id) %>%
  mutate(prev = if_else(value == 1, lag(value), 0L),
         prev = if_else(is.na(prev), 0L, prev)) %>%
  ungroup(id) %>%
  select(-value) %>%
  mutate(variable = "cont") %>%
  rename(value = prev)

n_set <- set_base$period2 %>% 
  add_region() %>% 
  group_by(period_label, reg) %>% 
  summarise(sum = n())

cont %>%
  filter(variable == "cont") %>%
  add_period_label() %>%
  add_region() %>%
  group_by(period_label, reg) %>%
  summarise(n = sum(value, na.rm = TRUE)) %>%
  full_join(n_set) %>%
  mutate(n = if_else(is.na(n), 0L, n),
         perc = (n / sum) * 100,
         reg = if_else(reg == "B", "East Bohemia", "Morava river catchment")) %>%
  ggplot(aes(period_label, perc)) +
  geom_col(color = "black", fill = "white", 
           position = "dodge", show.legend = FALSE) +
  facet_wrap(vars(reg), ncol = 1) +
  labs(x = "Period", y = "Continuity of settlements (%)") +
  theme_bw()

# ggsave(here::here("groups_continuity.svg"), width = 10, height = 5)
```

@@

```{r pca-plot}
pca_per %>% plot_nested_pca(var = "period")
# ggsave(here::here("analysis/groups/pca.pdf"), width = 21, height = 35)
```

```{r pca-plot2}
pca_per %>% plot_nested_pca(var = "period", pc = c("PC3", "PC4"))
# ggsave(here::here("analysis/groups/pca2.pdf"), width = 21, height = 35)
```
